---
title: "eAE Growth Analysis"
author: "Patrick Kiel"
date: "2023-12-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)

# load libraries
library(tidyverse)
library(seacarb)
library(lme4)
library(emmeans)
library(lsmeans)
library(lmerTest)
library(rstatix)
library(wesanderson)
library(ggpubr)
library(grid)
```

```{r, load data}
assignment <- read_csv("../data/acerExperiment/tankAssignment.csv") %>%
    mutate(tag = as.character(tag),
           tank = as.factor(tank))

bw <- read_csv("../data/acerExperiment/bwData.csv") %>%
        # sw calculate density from YSI vlaues
  mutate(rho = rho(S= sal, T = temp)/1000, .after="mass",
         # calculate mass in air from sw density, aragonite density, bw
          mass_a = mass/(1-rho/2.93),
         # convert date to date object
         date = as.Date(date, format="%m/%d/%Y"))

sa <- read_csv('../data/acerExperiment/surfaceAreas.csv') %>%
  mutate(tag = as.character(tag))

growth <- bw %>%
  left_join(assignment, by="tag") %>%
  left_join(sa, by = "tag") %>%
  arrange(tag,date) %>%
  group_by(tag) %>%
  # calculate mass gain in mg
  mutate(dailyMassGain = (mass_a - lag(mass_a))*1000/as.numeric(date-lag(date)),
         G = dailyMassGain/SA.cm2,
         days = as.numeric(date - first(date)),
         .after = "mass_a") %>%
  drop_na(tank, dailyMassGain)
```

```{r lmExtractFunction, include=F}
# function to extract linear model parameters

lmExtractCI <- function(mod) {
  combinedCoefs = c(coef(mod), confint(mod, level=0.95))
  names(combinedCoefs) = c("mean", "lowerCI", "upperCI")
  return(combinedCoefs)
}

```

# Controls

## How did control acrylic corals grow?

```{r acrylic-growth}
growth %>%
  filter(substrate=="inert") %>%
  ggplot(aes(x=date, y=dailyMassGain,
             group=interaction(as.factor(date),tank),
             fill = tank)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Daily Mass gain (mg/day)",
         title = "Acrylic Controls Daily Mass Gain")

growth %>%
  filter(substrate=="inert") %>%
  ggplot(aes(x=tank, y=dailyMassGain,
             fill = tank)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Daily Mass gain (mg/day)",
         title = "Acrylic Controls Daily Mass Gain")

growth %>%
  filter(substrate=="inert") %>%
  ggplot(aes(x=tank, y=G,
             fill = tank)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Daily Growth Rate (mg/day/cm2)",
         title = "Acrylic Controls Standardized Daily Growth Rate")

growth %>%
  filter(substrate=="inert") %>%
  group_by(substrate) %>%
  summarise(G_mean = mean(dailyMassGain, na.rm=T),
            G_std = sd(dailyMassGain, na.rm=T))

growth %>%
  filter(substrate=="inert") %>%
  group_by(substrate) %>%
  summarise(G_mean = mean(G, na.rm=T),
            G_std = sd(G, na.rm=T))

# no differences among tank
growth %>%
  filter(substrate=="inert") %>%
  lmer(dailyMassGain ~ tank + (1|tag), data=.) -> mod
anova(mod)
marginal = emmeans(mod, ~ tank)
pairs(marginal, adjust="tukey") %>%
  broom::tidy()

# no differences among genet
growth %>%
  filter(substrate=="inert") %>%
  lmer(dailyMassGain ~ genotype + (1|tag), data=.) -> mod
anova(mod)
marginal = emmeans(mod, ~ genotype)
pairs(marginal, adjust="tukey") %>%
  broom::tidy()
```

## How did the control eAE plugs abiotically grow?
```{r control-growth}
growth %>%
  filter(genotype=="control") %>%
  ggplot(aes(x=date, y=dailyMassGain,
             group=interaction(as.factor(date),tank),
             fill = tank)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Daily Mass gain (mg/day)",
         title = "All eAE Controls by Date")

# sometimes messy data if the glue fell off which was accounted for, remove those
growth %>%
  filter(genotype=="control" & (str_detect(notes, "glue", negate=T) | is.na(notes))) %>%
  ggplot(aes(x=date, y=dailyMassGain,
             group=interaction(as.factor(date),tank),
             fill = tank)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Daily Mass gain (mg/day)",
         title = "Filtered eAE Controls by Date")

growth %>%
  filter(genotype=="control" & (str_detect(notes, "glue", negate=T) | is.na(notes))) %>%
  ggplot(aes(x=tank, y=dailyMassGain,fill = tank)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Daily Mass gain (mg/day)",
         title = "Filtered eAE Controls by Tank")

growth %>%
  filter(genotype=="control" & (str_detect(notes, "glue", negate=T) | is.na(notes))) %>%
  group_by(genotype) %>%
 summarise(G_mean = mean(dailyMassGain, na.rm=T),
          G_std = sd(dailyMassGain, na.rm=T))

# no differences in abiotic growth per tank
growth %>%
  filter(genotype=="control" & (str_detect(notes, "glue", negate=T) | is.na(notes))) %>%
  lmer(dailyMassGain ~ tank + (1|tag), data=.) -> mod
shapiro.test(residuals(mod))
anova(mod)
marginal = emmeans(mod, ~ tank)
pairs(marginal, adjust="tukey") %>%
  broom::tidy()

# no differences in abiotic growth through time
growth %>%
  filter(genotype=="control" & (str_detect(notes, "glue", negate=T) | is.na(notes))) %>%
  lmer(dailyMassGain ~ days + (1|tag), data=.) -> mod
anova(mod)

growth %>%
  filter(genotype=="control") %>%
  filter(G > 0) %>%
  mutate(experiment = "acer") %>%
  write_csv(., '../data/microfragExperiment/abioticG_compare_acerExpData.csv')
```


# eAE Corals

We will first look at the total eAE growth and then subtract out this average eAE growth per tank for the corals to see how they perform against the combination of the bare cathode and acrylic controls. This is the basis of our investigation: do corals grown with eAE grow faster than controls after accounting for the abiotic precipitate.

## How did the eAE corals grow?

```{r eAE-corals}
growth %>%
  filter(substrate=="eAE" & genotype!="control") %>%
  ggplot(aes(date, dailyMassGain, group=interaction(as.factor(date),tank),
             fill=tank)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Daily Mass gain (mg/day)",
       title = "All eAE Corals by Tank and Date")

growth %>%
    filter(substrate=="eAE" & genotype!="control" & ((str_detect(notes, "glue", negate=T) & str_detect(notes, "electric", negate=T)) | is.na(notes))) %>%
  ggplot(aes(date, dailyMassGain, group=interaction(as.factor(date),tank),
             fill=tank)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Daily Mass gain (mg/day)",
       title = "Filtered eAE Corals by Tank and Date")

growth %>%
    filter(substrate=="eAE" & genotype!="control" & ((str_detect(notes, "glue", negate=T) & str_detect(notes, "electric", negate=T)) | is.na(notes))) %>%
  ggplot(aes(tank, dailyMassGain,
             fill=tank)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Daily Mass gain (mg/day)",
       title = "Filtered eAE Corals by Tank")

growth %>%
    filter(substrate=="eAE" & genotype!="control" & ((str_detect(notes, "glue", negate=T) & str_detect(notes, "electric", negate=T)) | is.na(notes))) %>%
  ggplot(aes(tank, massG,
             fill=tank)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Daily Growth Rates (mg/day/gram)",
       title = "eAE Corals' Standardized Daily Growth Rates by Tank")

growth %>%
   filter(substrate=="eAE" & genotype!="control" & ((str_detect(notes, "glue", negate=T) & str_detect(notes, "electric", negate=T)) | is.na(notes))) %>%
  group_by(substrate) %>%
  summarise(G_mean = mean(dailyMassGain, na.rm=T),
            G_sd = sd(dailyMassGain, na.rm=T))

# no diffeences in eAE by tank
growth %>%
    filter(substrate=="eAE" & genotype!="control" & ((str_detect(notes, "glue", negate=T) & str_detect(notes, "electric", negate=T)) | is.na(notes))) %>%
  aov(dailyMassGain ~ tank, data=.) %>%
  TukeyHSD() %>%
  broom::tidy()

# no differences in eAE by genet
growth %>%
    filter(substrate=="eAE" & genotype!="control" & ((str_detect(notes, "glue", negate=T) & str_detect(notes, "electric", negate=T)) | is.na(notes))) %>%
  aov(dailyMassGain ~ genotype, data=.) %>%
  TukeyHSD() %>%
  broom::tidy()
```

## Subtracting eAE controls from the eAE Corals

We can combine the means and confidence intervals for all the corals thus far to get a rough comparison to the eAE corals. I'm labeling this row "combined controls" below.

```{r}
# growth rates after substracting out mean abiotic growth
growth %>%
   filter(substrate=="eAE" & genotype!="control" & ((str_detect(notes, "glue", negate=T) & str_detect(notes, "electric", negate=T)) | is.na(notes))) %>%
  mutate(dailyMassGain = dailyMassGain - 34.04715) %>%
  group_by(substrate) %>%
  summarise(G_mean = mean(dailyMassGain, na.rm=T),
            G_sd = sd(dailyMassGain, na.rm=T))

growth %>%
   filter(substrate=="eAE" & genotype!="control" & ((str_detect(notes, "glue", negate=T) & str_detect(notes, "electric", negate=T)) | is.na(notes))) %>%
  mutate(G = (dailyMassGain - 34.04715)/SA.cm2) %>%
  group_by(substrate) %>%
  summarise(G_mean = mean(G, na.rm=T),
            G_sd = sd(G, na.rm=T))

## monte carlo error propagation methods ----
# monte carlo error propagation for the daily mass gain
set.seed(123)

# means & SD
bare_mean = 34.05
bare_sd = 6.69
combined_mean = 41.45
combined_sd = 3.96
  
# Number of simulations
n_sim <- 10000

# Generate random samples from each distribution
combinedMC <- rnorm(n_sim, mean = combined_mean, sd = combined_sd)
bareMC <- rnorm(n_sim, mean = bare_mean, sd = bare_sd)

# Subtract to get corrected coral growth
correctedMC <- combinedMC - bareMC

# Get summary stats
mean_corrected <- mean(correctedMC)
sd_corrected <- sd(correctedMC)

# Print results
cat("Corrected coral growth:", round(mean_corrected, 2), "±", round(sd_corrected, 2), "mg d^-1")
#-------


# create my new dataframe
growth_final <- growth %>%
   filter(substrate=="eAE" & genotype!="control" & ((str_detect(notes, "glue", negate=T) & str_detect(notes, "electric", negate=T)) | is.na(notes))) %>%
  mutate(substrate = "eAE Coral")
  
growth_final %>%
  bind_rows(growth_final %>%
  mutate(substrate = "eAE Coral Adjusted",
         dailyMassGain = dailyMassGain - 34.04715)) %>%
bind_rows(growth %>%
   filter(substrate=="inert") %>%
     mutate(substrate="Inert")) %>%
  bind_rows(
  growth %>%
  filter(substrate=="eAE" & genotype=="control" & ((str_detect(notes, "glue", negate=T) & str_detect(notes, "electric", negate=T)) | is.na(notes))) %>%
    mutate(substrate = "eAE Bare")) -> growth_final
  
# skip the blue for Moonrise3 palette
moonrise_colors <- c(wes_palette("Moonrise3")[3:5], wes_palette("Moonrise3")[2])
moonrise_colors_g <- c(wes_palette("Moonrise3")[5], wes_palette("Moonrise3")[2])


# Set factor levels
growth_final$substrate <- factor(growth_final$substrate,
                                 levels = c("eAE Coral", "eAE Bare",
                                            "eAE Coral Adjusted", "Inert"))
# adjust the G for the eAE Coral Adjusted Group
growth_final %>%
  filter(substrate == "Inert" | substrate == "eAE Coral Adjusted") %>%
  mutate(G = dailyMassGain/SA.cm2) -> growth_final_g


panel_a <- growth_final %>%
  ggplot(aes(substrate,dailyMassGain,fill=substrate)) +
  geom_boxplot(notch = F) +
  scale_fill_manual(values = moonrise_colors) +
  labs(x = "Substrate",
       #y = expression("Daily Growth Rate (mg "~day^-1*")"),
       y = bquote(bold("Daily Growth Rate (mg "~day^-1*")"))) +
  theme_bw() +
 theme(legend.position = "none",
        panel.border = element_rect(linewidth = 1),
        panel.grid = element_blank(),
        panel.background = element_rect(fill='transparent'), #transparent panel bg
        plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
        panel.grid.major = element_blank(), #remove major gridlines
        panel.grid.minor = element_blank(), #remove minor gridlines
        legend.background = element_rect(fill='transparent', color = NA, col=0), #transparent legend bg
        axis.text = element_text(size=10),
        axis.title = element_text(size=12, face="bold"),
        plot.caption.position = "plot",  # Moves caption inside the plot (bottom-left)
        plot.caption = element_text(hjust = 0, size = 10))  # Left-align & increase font size)

panel_b <- growth_final_g %>%
  ggplot(aes(substrate,G,fill=substrate)) +
  geom_boxplot(notch = F) +
  scale_fill_manual(values = moonrise_colors_g) +
  labs(x = "Substrate",
       y = bquote(bold("Daily Calcification Rate (mg "~cm^-2~day^-1*")"))) +
  theme_bw() +
 theme(legend.position = "none",
        panel.border = element_rect(linewidth = 1),
        panel.grid = element_blank(),
        panel.background = element_rect(fill='transparent'), #transparent panel bg
        plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
        panel.grid.major = element_blank(), #remove major gridlines
        panel.grid.minor = element_blank(), #remove minor gridlines
        legend.background = element_rect(fill='transparent', color = NA, col=0), #transparent legend bg
        axis.text = element_text(size=10),
        axis.title = element_text(size=12, face="bold"),
        plot.caption.position = "plot",  # Moves caption inside the plot (bottom-left)
        plot.caption = element_text(hjust = 0, size = 10))  # Left-align & increase font size)

# combine the plots
ggarrange(panel_a + rremove("xlab"), panel_b + rremove("xlab"), 
          labels = c("A", "B"),
          ncol = 2, nrow = 1,
          widths = c(3,2)) -> figure
annotate_figure(figure,
                bottom = text_grob("Substrate", face = "bold", size = 12))

ggsave('../figures/acerGrowthPlot.tiff',
       width = 9, height=5, dpi=600,
       units = "in",
       device = "tiff", compression ="lzw")

#stats
ggplot(growth_final_g, aes(sample = (dailyMassGain))) +
  stat_qq() + 
  stat_qq_line() +
  theme_minimal() +
  labs(title = "QQ Plots for Daily Growth Rate", x = "Theoretical Quantiles", y = "Sample Quantiles")

var.test(G~substrate, data = growth_final_g) # p > 0.05 can assume equal variance

# proceed with lmer to account for repeated measures
mod <- lmer(G ~ substrate + (1 | tank) + (1 | tag),
              data = growth_final_g)
summary(mod)
anova(mod)

```