---
title: "Planar Growth Analysis"
output: html_notebook
---

## Load packages
```{r load-packages}
library(tidyverse)
library(wesanderson)
library(lme4)
library(emmeans)
library(lsmeans)
library(lmerTest)
library(rstatix)
```

## Import Data
```{r import-data}
areas <- read_csv('../data/microfragExperiment/planarAreas.csv')
photoIDs <- read_csv('../data/microfragExperiment/photoIDs.csv')
tankAssignment <- read_csv('../data/microfragExperiment/tankAssignment.csv')
```


## Merge Data
```{r}
planarPhotoIds <- unique(areas$photoID)
vector_df <- tibble(photoID = planarPhotoIds)

photoIDs %>%
  pivot_longer(pos1:pos5,
               names_pattern = "pos(.*)",
               names_to = "position",
               values_to = "id",
               values_drop_na = T) %>%
  mutate(id_lower = as.integer(str_sub(photoIDs, 1,3)),
         id_upper = as.integer(str_sub(photoIDs, -3)),
         date = parse_date(date, format="%m/%d/%Y"),
         position = as.integer(position)) %>%
  select(-photoIDs) %>%
  rowwise() %>%
  mutate(photoID = list(vector_df$photoID[vector_df$photoID >= id_lower & vector_df$photoID <= id_upper])) %>%
  unnest(photoID) %>%
  select(-c(id_lower,id_upper)) -> photoIDs

# cast correct column types
areas <- areas %>%
  mutate(date = parse_date(date, format="%m/%d/%Y"))

photoIDs %>%
  left_join(areas,
            by = c("date",
                   "position",
                   "photoID")) -> planarAreas

planarAreas %>%
  left_join(tankAssignment, by ="id") %>%
  select(date, photoID, position, id, area, x, y, substrate:tank) -> planarAreas

rm(planarPhotoIds, vector_df)
```


## total percent growth data
```{r}
planarAreas %>%
  filter(height != "blank") %>%
  arrange(id, date) %>%
  group_by(id, substrate, genotype, height, tank) %>%
  summarise(per_growth = (last(area)-first(area))/first(area)) -> perGrowthRates
```

```{r}
perGrowthRates %>%
  group_by(substrate, height) %>%
  summarise(growthRate = mean(per_growth),
            growthrate_sd = sd(per_growth)) %>%
  ungroup()
```



## Emtrends slope technique to account for high variability

Splitting by short/tall for morphological differences as short corals can reach the substrate and begin skirting. Additionally  Then looking for differences in slope between the corals.

```{r}
moonrise_subset <- c(rep(c(wes_palette("Moonrise3")[5], wes_palette("Moonrise3")[2]),2))

planarAreas %>%
  filter(height!="blank") %>%
  mutate(sh = paste(substrate, height, sep="-"),
         height = str_to_title(height),
         substrate = case_when(substrate == "inert" ~ "Inert", T ~ substrate)) %>%
  ggplot(aes(date, area, shape = substrate,
             group=as.factor(sh))) +
  geom_point(size=3) +
  geom_smooth(method="lm", aes(group=substrate,
                               linetype = substrate),
              color="black",
              show.legend = F) +
  # duplicate to hack the legend
  geom_smooth(method="lm", aes(group=substrate,
                             linetype = substrate),
            color="black", se = F,
            show.legend = T) +
  scale_linetype_manual(values = c("eAE"=1,"Inert"=3)) +
  #scale_color_manual(values = moonrise_subset) +
  scale_x_date(limits = c(as.Date("2024-02-15"), as.Date("2024-05-01"))) +
  facet_wrap(~height) +
  labs(shape = "Substrate", linetype = "Substrate",
       x = "Date", y= bquote(bold('Area '~(cm^2)))) +
  # geom_text(data = data.frame(x = as.Date("2024-03-01"), area = Inf,
  #                             #labs = c("Short", "Tall"),
  #                             height = c("Short", "Tall")),
  #        aes(x = x, y=area, label = height), inherit.aes = F,
  #        hjust = 0, vjust = 1.6,
  #        nudge_x = -15,
  #        size = 5, fontface = "bold") +
  #guides(linetype = )
  theme_bw() +
  theme(#legend.position = "bottom",
        #strip.background = element_blank(),
        legend.position=c(0.14, 0.87),
        legend.direction = "horizontal",
        legend.title.position = "top",
        legend.title = element_text(size=10, face="bold"),
        legend.box.background = element_rect(color="black", linewidth = 0.5),
        strip.text.x = element_text(size = 14, face="bold"),
        panel.border = element_rect(linewidth = 1),
        panel.grid = element_blank(),
        panel.background = element_rect(fill='transparent'), #transparent panel bg
        plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
        panel.grid.major = element_blank(), #remove major gridlines
        panel.grid.minor = element_blank(), #remove minor gridlines
        legend.background = element_rect(fill='transparent', color = NA, col=0), #transparent legend bg
        axis.text = element_text(size=12),
        axis.title = element_text(size=14, face="bold"),
        plot.caption.position = "plot",  # Moves caption inside the plot (bottom-left)
        plot.caption = element_text(hjust = 0, size = 10))  # Left-align & increase font size)

ggsave('../figures/planarGrowth_CI.svg',
       dpi = 600,
       width = 10,
       height = 5,
       units = "in",
       device = "svg")
       #compression = "lzw")


planarAreas %>%
  filter(height!="blank") %>%
  group_by(substrate,height) %>%
  nest() %>%
  mutate(slope = map(data, function(df) {
                        mod = lm(area ~ date, data=df)
                        return(coefficients(mod)[2])
                        })) %>%
  unnest(slope) %>%
  arrange(substrate) %>%
  group_by(height) %>%
  mutate(ratio = slope/lag(slope))

```

## stats
```{r}
planarAreas %>%
  arrange(id, date) %>%
  group_by(id) %>%
  mutate(days = as.numeric(date - first(date))) %>%
  ungroup() -> planarAreas

model <- lmer(area ~ days * substrate * height + (1|tank) + (1 | id), data=planarAreas)

# check model assumptions
ggplot(planarAreas, aes(sample = (area))) +
  stat_qq() + 
  stat_qq_line() +
  theme_minimal() +
  labs(title = "QQ Plots for Planar Areas", x = "Theoretical Quantiles", y = "Sample Quantiles")

shapiro.test(residuals(model)) # residuals are normal, agrees w/ QQQ
boxplot(residuals(model) ~ planarAreas$height) # homoscedascity 

# Model summary
summary(model)

# Estimated marginal means slopes for pairwise comparisons
growth_rates <- emtrends(model, ~ substrate * height, var = "days")

# Pairwise comparisons of slopes
pairwise_comparisons <- pairs(growth_rates, adjust="tukey")

# Display results
growth_rates %>%
  broom::tidy() %>%
  mutate(SD = std.error * sqrt(df),
         .after= "days.trend")
pairwise_comparisons

percent_diff_ci <- function(slope1, se1, slope2, se2, df, conf=0.95) {
  diff <- slope1 - slope2
  percent_diff <- (diff / slope2) * 100
  
  # Standard error of percent difference
  se_diff <- sqrt(se1^2 + se2^2)
  se_percent <- (se_diff / abs(slope2)) * 100
  
  # t-value for the given confidence level
  t_val <- qt(1 - (1 - conf)/2, df)
  
  # Confidence interval
  lower <- percent_diff - t_val * se_percent
  upper <- percent_diff + t_val * se_percent
  
  c(percent_diff = percent_diff, lower_CI = lower, upper_CI = upper)
}

broom::tidy(growth_rates) -> growth_rates
bind_cols(
comparison = c("eAE short vs inert short",
               "eAE short vs eAE tall",
               "eAE short vs inert tall"),

#eAE short vs inert short
bind_rows(
percent_diff_ci(growth_rates$days.trend[2],growth_rates$std.error[2],
                growth_rates$days.trend[1],growth_rates$std.error[1],
                df=137),

#eAE short vs eAE tall
percent_diff_ci(growth_rates$days.trend[2],growth_rates$std.error[2],
                growth_rates$days.trend[4],growth_rates$std.error[4],
                df=137),

# eAE short vs inert tall
percent_diff_ci(growth_rates$days.trend[2],growth_rates$std.error[2],
                growth_rates$days.trend[3],growth_rates$std.error[3],
                df=137)
))
```

Any differences among genotypes?

Can I remove from the random effects?


## Assess Errors

Checking errors by looking at the consistency of substrate area

```{r}
areaErrors <- readxl::read_excel('../data/microfragExperiment/photoScaleCheck.xlsx')

areaErrors %>%
  group_by(substrate) %>%
  summarise(mean = mean(diameter, na.rm = T),
          sd = sd(diameter, na.rm=T)) %>%
  mutate(RSD = 100*sd/mean) -> areaErrors
```
The average acrylic substrate was measured to be `r round(areaErrors$mean[1],3)` $\pm$ `r round(areaErrors$sd[1],3)` $cm^2$, granting a relative standard deviation (RSD) of `r round(areaErrors$RSD[1],3)`%. The average steel substrate was measured to be `r round(areaErrors$mean[2],3)` $\pm$ `r round(areaErrors$sd[2],3)` $cm^2$, granting a relative standard deviation (RSD) of `r round(areaErrors$RSD[2],3)`%. These random errors represent less than `r round(max(areaErrors$RSD))`% of the target measurements throughout the monitoring period. Therefore, the scaled photographic measuring technique was deemed acceptable for this analysis. 