---
title: "Microprofile Analysis and Graphs"
output: html_notebook
---

```{r, load-libraries}
library(tidyverse)
library(seacarb)
library(minpack.lm)

```

```{r, load-data}
jProfiles <- read_csv("../data/pHProfiles/currentDensityProfiles.csv")
flowProfiles <- read_csv("../data/pHProfiles/flowProfiles.csv")
# J = current density in A/m2
# v = flume bulk velocity in cm/s
# substrate = inert (control) / eAE 
# replicate = 3 runs per profile
# z = height above substrate in microns
# pH_T_avg = average station reading recentered to bulk values (pH=8.02), measured in pHnbs and converted to total scale using seacarb
# H_avg = pH_T_avg (total scale) converted to [H+] in nM
```

# Current Density Profiles

## Plot

```{r}
jProfiles %>%
  mutate(J = case_when(substrate == "inert" ~ 0,
                       T ~ J),
         substrate = as.factor(substrate)) -> jProfiles
jProfiles$J = factor(jProfiles$J, levels = c("0", "0.5", "1", "3"))
jProfiles %>%  
  ggplot(aes(z, pH_T_avg, color=J, shape = substrate)) +
  stat_summary(
    fun = mean, 
    fun.min = \(x) mean(x) - sd(x), 
    fun.max = \(x) mean(x) + sd(x),
    geom = "pointrange") +
  stat_summary(
    fun = mean,
    geom = "line"
  ) +
  geom_vline(xintercept = c(10305.15, 14644.82,  21189.76), linetype="dashed",
             color = c("#66C2A5","#FC8D62","#8DA0CB"), linewidth=0.75) + #dbl heights calculated below
  annotate(geom="rect",
    xmin = c(10305.15 - 823.2307, 14644.82 - 848.4919,  21189.76 - 676.8799),   
    xmax = c(10305.15 + 823.2307, 14644.82 + 848.4919,  21189.76 + 676.8799), 
    ymin = -Inf,
    ymax = Inf,
    fill = c("#66C2A5","#FC8D62","#8DA0CB"),
    alpha=0.3) +
  theme_bw() +
  theme_bw() +
  labs(x = "Height (mm)",
       y = bquote(bold(pH[T])),
       color = expression("Current Density, " ~ J ~ "("*A~m^-2*")"),
       shape = "Substrate") +
  scale_color_manual(values =c("0"="#000",
                               "0.5"="#66C2A5",
                               "1"="#FC8D62",
                               "3"="#8DA0CB"),
                     breaks = c("0", "0.5", "1", "3")) + 
  guides(shape = guide_legend(override.aes = list(linetype = 0)),
         color = guide_legend(breaks = c("0", "0.5", "1", "3"))) +
  scale_x_continuous(labels = function(x) x / 1000) +
theme(legend.position = "bottom",
        panel.border = element_rect(linewidth = 1),
        panel.grid = element_blank(),
        panel.background = element_rect(fill='transparent'), #transparent panel bg
        plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
        panel.grid.major = element_blank(), #remove major gridlines
        panel.grid.minor = element_blank(), #remove minor gridlines
        legend.background = element_rect(fill='transparent', color = NA, col=0), #transparent legend bg
        axis.text = element_text(size=10),
        axis.title = element_text(size=12, face="bold"),
        plot.caption.position = "plot",  # Moves caption inside the plot (bottom-left)
        plot.caption = element_text(hjust = 0, size = 10))  # Left-align & increase font size)

ggsave('../figures/currentDensityBoundaryLayer_alt.tiff',
       dpi = 600,
       width = 8,
       height = 4,
       units = "in",
       device = "tiff",
       compression = "lzw")
```

## Boundary Layer Calculations
3 methods:
1) 99% of H+
2) <10% different 4 consecutive times
3) Intersection of profile w/ hyperbolic tangent model

**Only method #3 used for manuscript**

```{r}
# function for dbl calc 2
# consecutive_method <- function(vec, z, threshold = 0.1, n = 4) {
#     # vec = vector of numbers iterated over
#     # threshold = % difference between neighboring values
#     # n = required number of consecutive values below the threshold
#   
#   # make all the pH > 8.25 NA since very rapid decrease then plateud changes in pH
#   # vec = ifelse(vec < 10^(-8.25), NA, vec)
#     
#   changes = abs(vec - lag(vec)) / lag(vec)  # percentage change
#   within_threshold = changes <= threshold  # % change <= threshold
# 
#   # Find the first sequence of > n consecutive TRUEs
#   consec_count = rle(within_threshold)  # Run-length encoding
#   valid_indices = which(consec_count$values & consec_count$lengths >= n)
#   
#     if (length(valid_indices) > 0) {
#     return(z[sum(consec_count$lengths[1:(valid_indices[1]-1)])+4]) 
#     } else {
#     return(NA) # Return NA if no valid sequence found
#   }
# }

# hyberbolic tangent intersection method, adapted from Nishihara and Ackerman (2007)
htanIntersecton <- function(dat, pH_filt=8.25) {
  dat = dat %>%
    mutate(z_norm = (z-min(z))/(max(z)-min(z))) %>% # scale z between 0 to 1
    filter(pH_T_avg < pH_filt) # remove the initial steps w/ rapidly decreasing pH to improve model fit

# Define the hyperbolic tangent model function  
  # tanh_model = function(par, depth) {
  #   A = par[1]  # Scaling factor
  #   B = par[2]  # Transition depth
  #   C = par[3]  # Steepness
  #   D = par[4]  # Offset
  #   A * tanh(C * (depth - B)) + D
  # }

# Initial parameter estimates (adjust as needed)
start_params = c(A = max(dat$H_avg)-min(dat$H_avg),
                 B = dat$z_norm[which.min(abs(dat$H_avg - median(dat$H_avg)))],
                 C = 0.1,
                 D = min(dat$H_avg))

# Fit the model
fit <- tryCatch({
      nlsLM(H_avg ~ A * tanh(C * (z_norm - B)) + D,
            data = dat,
            start = as.list(start_params),
            control = list(maxiter = 1000))
    }, 
    error = function(e) {
      profile_name <- if('v' %in% names(dat)) {
        paste0("Velocity = ", unique(dat$v), "; J = ", unique(dat$J), "; replicate = ", unique(dat$replicate))
      } else {
        paste0("J = ", unique(dat$J), "; replicate = ?")
      }
      message("Fit failed for profile: ", profile_name)
      message("Error: ", e$message)
      return(NULL)
    })

if(!is.null(fit)) {
# Extract fitted parameters
params = coef(fit)

# Generate fitted values for intersection
dat$fitted = params[1] * tanh(params[3] * (dat$z_norm - params[2])) + params[4]

# Identify the boundary layer height (where fitted profile meets the constant part)
asymptote = mean(tail(dat$fitted, n = 2))  # Assumed bulk concentration
threshold = asymptote * 0.99 # set limit to 99% of this bulk concentration 
boundary_height_data = max(dat$z[dat$fitted <= threshold])  # extract max z below threshold

# modeled height
boundary_height_model = data.frame(
  z_norm = seq(0,1,length.out=2000) # generates 2000 even steps across profile
) %>%
  mutate(z= z_norm*25000, # convert back to um for a 2.5cm profile
         modeled_values = params[1] * tanh(params[3] * (z_norm - params[2])) + params[4]) %>% # calculate the fitted values
  filter(modeled_values <=threshold) %>%
  pull(z) %>% max() # identify boundary layer before constant part of model

return(data.frame(dbl_model = boundary_height_model,
         dbl_data = boundary_height_data))
}
else{
  return(c(NA,
           NA))
}
}

# analyzeProfile <- function(profile) {
#   # 99% method
#   # grab last 5 [H+], calculate when profile hits 99% of this
#   # target_99 = mean(tail(profile$H_avg),na.rm = T)*0.99
#   # dbl_99 = profile$z[which.max(profile$H_avg[profile$H_avg < target_99])]
#   # 
#   # # <10% different 4 consecutive times
#   # dbl_consecutive = consecutive_method(vec = profile$H_avg,
#   #                                      z = profile$z,
#   #                                      threshold = 0.1)
#   
#   # hyperbolic tangent method
#   dbl_htan = htanIntersecton(profile)
#   
#   return(tibble(#dbl_99 = dbl_99,
#                  #dbl_consecutive = dbl_consecutive,
#                  dbl_htanmodel = dbl_htan[1],
#                  dbl_htandata = dbl_htan[2]))
# }

jProfiles %>%
  filter(substrate=="eAE") %>%
  group_by(date, J, substrate, replicate) %>%
  nest() %>%
  mutate(dbl = map(data, ~htanIntersecton(.x))) %>%
  unnest(dbl) %>%
  pivot_longer(cols = dbl_model:dbl_data,
               names_prefix = "dbl_",
               names_to = "method",
               values_to = "dbl")  -> jProfiles_dbl
jProfiles_dbl %>%
  group_by(J, method) %>%
  filter(method == "model") %>%
  summarise(mean = mean(dbl, na.rm = T),
            sd = sd(dbl, na.rm =T),
            n = n(),
            se = sd(dbl) / sqrt(n()),  
            ci_lower = mean(dbl) - qt(0.975, df = n() - 1) * se,
            ci_upper = mean(dbl) + qt(0.975, df = n() - 1) * se) %>%
  ungroup() -> j_dbl
# j_dbl %>%
#   filter(method =="htanmodel") %>%
#   write.table("clipboard",sep="\t",row.names=F)


jProfiles_dbl %>%
  filter(method == "model") -> j_dbl

j_dbl %>%
  aov(dbl ~ as.factor(J), data=.) -> j_aov

# **1. Check ANOVA Assumptions**
# Normality assumption - Shapiro-Wilk test per group
shapiro.test(residuals(j_aov))

# Homogeneity of variance - Levene's test
car::leveneTest(pH_T_avg ~ as.factor(J),
                jProfiles %>%
  filter(z==0 & substrate== "eAE"))

# **2. Run ANOVA**
j_aov %>%
  summary()

# **3. Post hoc test (Tukey HSD)**
TukeyHSD(j_aov)


jProfiles_dbl %>%
  # mutate(method = fct_relevel(method, 
  #                             "99", "htandata", "htanmodel", "consecutive")) %>%
  filter(method == "model" & substrate == "eAE") %>%
  ggplot(aes(x= as.factor(J), y= dbl,
             fill = as.factor(J), color = method)) +
  #geom_boxplot() +
  geom_violin() +
  scale_fill_brewer(palette = "Set2") +
  scale_color_manual(values = c("99"="black",
                                "consecutive"="black",
                                "htanmodel"="black",
                                "htandata"="black")) +
  theme_bw() +
  labs(y = "Boundary Layer Height") +
  scale_y_continuous(labels = function(y) y / 1000) +
  labs(x="Current Density",
       fill = expression("Current Density, " ~ J ~ "(" ~ A/m^2 ~ ")"),
       color = "DBL Method, L to R") +
  theme(legend.position = "none",
        panel.border = element_rect(linewidth = 1),
        panel.grid = element_blank(),
        axis.text = element_text(size=14),
        axis.title = element_text(size=16),
        plot.caption.position = "plot",  # Moves caption inside the plot (bottom-left)
        plot.caption = element_text(hjust = 0, size = 10))

# ggsave('../figures/currentDensityBoundaryHeight.tiff',
#        dpi = 600,
#        width = 5,
#        height = 5,
#        units = "in",
#        device = "tiff",
#        compression = "lzw")
```


## pH at certain heights
```{r}
jProfiles %>%
  filter(z==0 & substrate== "eAE") %>%
  group_by(J) %>%
  summarise(pH_avg = mean(pH_T_avg),
            pH_sd = sd(pH_T_avg))

jProfiles %>%
  filter(z==0 & substrate== "eAE") %>%
  aov(pH_T_avg ~ as.factor(J), data=.) -> j_aov

# **1. Check ANOVA Assumptions**
# Normality assumption - Shapiro-Wilk test per group
shapiro.test(residuals(j_aov))

# Homogeneity of variance - Levene's test
car::leveneTest(pH_T_avg ~ as.factor(J),
                jProfiles %>%
  filter(z==0 & substrate== "eAE"))

# **2. Run ANOVA**
j_aov %>%
  summary()

# **3. Post hoc test (Tukey HSD)**
TukeyHSD(j_aov)

# stats @ 5mm
jProfiles %>%
  filter(z == 5000 & substrate=="eAE") %>%
  group_by(J) %>%
  summarise(pH = mean(pH_T_avg),
            sd = sd(pH_T_avg, na.rm =T))

jProfiles %>%
  filter(z == 5000 & substrate=="eAE") %>%
  aov(pH_T_avg ~ as.factor(J), data=.) -> j_aov

# **1. Check ANOVA Assumptions**
# Normality assumption - Shapiro-Wilk test per group
shapiro.test(residuals(j_aov))

# Homogeneity of variance - Levene's test
car::leveneTest(pH_T_avg ~ as.factor(J),
                jProfiles %>%
                  filter(z == 5000 & substrate=="eAE"))

# **2. Run ANOVA**
j_aov %>%
  summary()

# **3. Post hoc test (Tukey HSD)**
TukeyHSD(j_aov)

# stats @ 15mm
jProfiles %>%
  filter(z == 15000 & substrate=="eAE") %>%
  group_by(J) %>%
  summarise(pH = mean(pH_T_avg),
            sd = sd(pH_T_avg, na.rm =T))

jProfiles %>%
  filter(z == 15000 & substrate=="eAE") %>%
  aov(pH_T_avg ~ as.factor(J), data=.) -> j_aov

# **1. Check ANOVA Assumptions**
# Normality assumption - Shapiro-Wilk test per group
shapiro.test(residuals(j_aov))

# Homogeneity of variance - Levene's test
car::leveneTest(pH_T_avg ~ as.factor(J),
                jProfiles %>%
                  filter(z == 15000 & substrate=="eAE"))

# **2. Run ANOVA**
j_aov %>%
  summary()

# **3. Post hoc test (Tukey HSD)**
TukeyHSD(j_aov)

```

## stats
```{r}
jProfiles_dbl %>%
  filter(method == "model") -> jDBL_htanmodel

# **1. Check ANOVA Assumptions**
# Normality assumption - Shapiro-Wilk test per group
jDBL_htanmodel %>%
  group_by(J) %>%
  summarise(shapiro_p = shapiro.test(dbl)$p.value)

# Homogeneity of variance - Levene's test
  car::leveneTest(dbl ~ as.factor(J), jDBL_htanmodel)

# **2. Run ANOVA**
aov(dbl ~ as.factor(J), jDBL_htanmodel) -> jModel
jModel %>%
  summary()


# **3. Post hoc test (Tukey HSD)**
TukeyHSD(jModel)
```



## BL Approximate to Scaling Laws?

Check if the measured dbl is within the approximate scaling of each other

```{r}
data.frame(J = c(0.5, 1, 3),
           dbl_actual = subset(j_dbl, method=="htanmodel")$mean) %>%
          mutate(dbl_model = (J/1)^(1/3)*subset(j_dbl, method=="htanmodel")$mean[2])

```


# Flow Profiles

## Plot
```{r}
flowProfiles %>%
  mutate(v = as.factor(v),
         substrate = as.factor(substrate)) -> flowProfiles
flowProfiles %>%  
  ggplot(aes(z, pH_T_avg, color=v)) +
  stat_summary(
    fun = mean, 
    fun.min = \(x) mean(x) - sd(x), 
    fun.max = \(x) mean(x) + sd(x),
    geom = "pointrange") +
  stat_summary(
    fun = mean,
    geom = "line"
  ) +
  geom_vline(xintercept = c(5000,15000), linetype="dotted") + # microfrag heights
  geom_vline(xintercept = c(21344.005, 15049.191,  5919.626), linetype="dashed",
             color = c("#66C2A5","#FC8D62","#8DA0CB"), linewidth=0.75) + #dbl heights calculated below
  annotate(geom="rect",
    xmin = c(21344.005 - 229.2434, 15049.191 - 1819.62043,  5919.626 - 1438.50908),   
    xmax = c(21344.005 + 229.2434, 15049.191 + 1819.62043,  5919.626 + 1438.50908), 
    ymin = -Inf,
    ymax = Inf,
    fill = c("#66C2A5","#FC8D62","#8DA0CB"),
    alpha=0.3) +
  theme_bw() +
  labs(x = "Height (mm)",
       y = bquote(bold(pH[T])),
       color = expression("Bulk Flow, " ~ v ~ "("*cm~s^-1*")"),
       shape = "Substrate") +
  scale_color_brewer(palette = "Set2") + 
  guides(shape = guide_legend(override.aes = list(linetype = 0))) +
  scale_x_continuous(labels = function(x) x / 1000) +
  theme(legend.position = "bottom",
        panel.border = element_rect(linewidth = 1),
        panel.grid = element_blank(),
        panel.background = element_rect(fill='transparent'), #transparent panel bg
        plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
        panel.grid.major = element_blank(), #remove major gridlines
        panel.grid.minor = element_blank(), #remove minor gridlines
        legend.background = element_rect(fill='transparent', color = NA, col=0), #transparent legend bg
        axis.text = element_text(size=10),
        axis.title = element_text(size=12, face="bold"),
        plot.caption.position = "plot",  # Moves caption inside the plot (bottom-left)
        plot.caption = element_text(hjust = 0, size = 10))  # Left-align & increase font size)

ggsave('../figures/bulkFlowBoundaryLayer_alt.tiff',
       dpi = 600,
       width = 8,
       height = 4,
       units = "in",
       device = "tiff",
       compression = "lzw")
```

## pH at certain heights
```{r}
flowProfiles %>%
  filter(z==0) %>%
  summarise(pH_avg = mean(pH_T_avg),
            pH_sd = sd(pH_T_avg))
```

## Boundary Layer Calculations
```{r}
flowProfiles %>%
  group_by(date, v,  replicate) %>%
  nest() %>%
  mutate(dbl = map(data, ~htanIntersecton(.x))) %>%
  unnest(dbl) %>%
  pivot_longer(cols = dbl_model:dbl_data,
               names_prefix = "dbl_",
               names_to = "method",
               values_to = "dbl") -> flowProfiles_dbl

flowProfiles_dbl %>%
  group_by(v, method) %>%
  summarise(mean = mean(dbl, na.rm = T),
            sd = sd(dbl, na.rm =T),
            n = n(),
            se = sd(dbl) / sqrt(n()),  
            ci_lower = mean(dbl) - qt(0.975, df = n() - 1) * se,
            ci_upper = mean(dbl) + qt(0.975, df = n() - 1) * se) %>%
  ungroup() -> flow_dbl

# flow_dbl %>%
#   filter(method =="htanmodel") %>%
#   write.table("clipboard",sep="\t",row.names=F)

flowProfiles_dbl %>%
  # mutate(method = fct_relevel(method, 
  #                             "99", "htandata", "htanmodel", "consecutive")) %>%
  filter(method == "model") %>%
  ggplot(aes(x= as.factor(v), y= dbl,
             fill = as.factor(v))) +
  #geom_boxplot() +
  geom_violin() +
  scale_fill_brewer(palette = "Set2") +
  # scale_color_manual(values = c("99"="black",
  #                               "consecutive"="black",
  #                               "htanmodel"="black",
  #                               "htandata"="black")) +
  theme_bw() +
  labs(y = "Boundary Layer Height") +
  scale_y_continuous(labels = function(y) y / 1000) +
  labs(x="Bulk Flow, v (cm/s)",
       fill ="Bulk Flow, v (cm/s)",
       color = "DBL Method, L to R") +
theme(legend.position = "none",
        panel.border = element_rect(linewidth = 1),
        panel.grid = element_blank(),
        panel.background = element_rect(fill='transparent'), #transparent panel bg
        plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
        panel.grid.major = element_blank(), #remove major gridlines
        panel.grid.minor = element_blank(), #remove minor gridlines
        legend.background = element_rect(fill='transparent', color = NA, col=0), #transparent legend bg
        axis.text = element_text(size=10),
        axis.title = element_text(size=12, face="bold"),
        plot.caption.position = "plot",  # Moves caption inside the plot (bottom-left)
        plot.caption = element_text(hjust = 0, size = 10))  # Left-align & increase font size)

# ggsave('../figures/BulkFlowBoundaryHeight.tiff',
#        dpi = 600,
#        width = 5,
#        height = 5,
#        units = "in",
#        device = "tiff",
#        compression = "lzw")

```

```{r}
flowProfiles_dbl %>%
  filter(v==1) %>%
  group_by(v) %>%
  summarise(dbl_mean = mean(dbl),
            dbl_sd = sd(dbl))

flowProfiles_dbl %>%
  filter(method == "model") -> flow_dbl

flow_dbl %>%
  group_by(v) %>%
  summarise(dbl_avg = mean(dbl),
            dbl_sd = sd(dbl)) %>%
  mutate(across(c(dbl_avg, dbl_sd), ~.x/1000))

flow_dbl %>%
  aov(dbl ~ as.factor(v), data=.) -> flow_aov

# **1. Check ANOVA Assumptions**
# Normality assumption - Shapiro-Wilk test per group
shapiro.test(residuals(flow_aov))

# Homogeneity of variance - Levene's test
car::leveneTest(dbl ~ as.factor(v), flow_dbl)

# **2. Run ANOVA**
flow_aov %>%
  summary()

# **3. Post hoc test (Tukey HSD)**
TukeyHSD(flow_aov)
```

```{r}
# stats @ 0mm
flowProfiles %>%
  filter(z == 0) %>%
  group_by(v) %>%
  summarise(pH = mean(pH_T_avg),
            sd = sd(pH_T_avg, na.rm =T))

flowProfiles %>%
  filter(z == 0) %>%
  aov(pH_T_avg ~ as.factor(v), data=.) -> flow_aov

# **1. Check ANOVA Assumptions**
# Normality assumption - Shapiro-Wilk test per group
shapiro.test(residuals(flow_aov))

# Homogeneity of variance - Levene's test
car::leveneTest(pH_T_avg ~ as.factor(v), flowProfiles %>%
  filter(z == 5000))

# **2. Run ANOVA**
flow_aov %>%
  summary()

# **3. Post hoc test (Tukey HSD)**
TukeyHSD(flow_aov)
# stats @ 5mm
flowProfiles %>%
  filter(z == 5000) %>%
  group_by(v) %>%
  summarise(pH = mean(pH_T_avg),
            sd = sd(pH_T_avg, na.rm =T))

flowProfiles %>%
  filter(z == 5000) %>%
  aov(pH_T_avg ~ as.factor(v), data=.) -> flow_aov

# **1. Check ANOVA Assumptions**
# Normality assumption - Shapiro-Wilk test per group
shapiro.test(residuals(flow_aov))

# Homogeneity of variance - Levene's test
car::leveneTest(pH_T_avg ~ as.factor(v), flowProfiles %>%
  filter(z == 5000))

# **2. Run ANOVA**
flow_aov %>%
  summary()

# **3. Post hoc test (Tukey HSD)**
TukeyHSD(flow_aov)

# stats @ 15mm
flowProfiles %>%
  filter(z == 15000) %>%
  group_by(v) %>%
  summarise(pH = mean(pH_T_avg),
            sd = sd(pH_T_avg, na.rm =T))

flowProfiles %>%
  filter(z == 15000) %>%
  aov(pH_T_avg ~ as.factor(v), data=.) -> flow_aov

# **1. Check ANOVA Assumptions**
# Normality assumption - Shapiro-Wilk test per group
shapiro.test(residuals(flow_aov))

# Homogeneity of variance - Levene's test
car::leveneTest(pH_T_avg ~ as.factor(v), flowProfiles %>%
  filter(z == 15000))

# **2. Run ANOVA**
flow_aov %>%
  summary()

# **3. Post hoc test (Tukey HSD)**
TukeyHSD(flow_aov)

```


## What is the pH @ heights of coral?

```{r}
flowProfiles %>%
  filter(z == 5000 | z== 15000) %>%
  bind_rows(
    jProfiles %>%
      filter(z == 5000 | z== 15000) %>%
      filter(substrate=="eAE") %>%
      mutate(v=1)
  ) %>%
  group_by(v, z, J) %>%
  summarise(pH = mean(pH_T_avg),
            sd = sd(pH_T_avg, na.rm =T),
            n = n(),
            se = sd(pH_T_avg) / sqrt(n()),  
            ci_lower = mean(pH_T_avg) - qt(0.975, df = n() - 1) * se,
            ci_upper = mean(pH_T_avg) + qt(0.975, df = n() - 1) * se) %>%
  mutate(across(c(pH, sd, se, ci_lower, ci_upper), ~signif(.x,4)),
         z = z/1000) %>%
  arrange(v, J)#%>%
  #write.table("clipboard",sep="\t",row.names=F)
  
```

## Model Omega @ heights of coral
```{r}
## avg measured values across experiments
dic_init <- 2108e-6
ta_init <- 2436e-6
S <- 34
Temp <- 27
pHFinal = flowProfiles %>%
  filter((z == 5000 | z== 15000) & v == 1 ) %>%
  group_by(z) %>%
  summarise(pH = mean(pH_T_avg)) %>%
  pull(pH)

# pHFinal<-8.16

# Function to calculate carb system after adding carbonate
carbAddition <- function(dic_init, ta_init, species, mineral_addition, S, T) {
  if(species == "co3") {
    dic_new <- dic_init + mineral_addition
    ta_new <- ta_init + 2*mineral_addition
  } else if(species == "hco3") {
    dic_new <- dic_init + mineral_addition
    ta_new <- ta_init + mineral_addition
  } else if(species == "oh") {
    dic_new <- dic_init
    ta_new <- ta_init + mineral_addition
  } else{
    return(NA)
  }
  # Use carb function to calculate pH from DIC and TA
  # Flag 15 means we're providing ALK and DIC as inputs
  result <- carb(flag = 15, var1 = ta_new, var2 = dic_new, 
                 S = S, T = Temp)
  
  return(result)
}

# Find what co3 added would get us to pH
CO3_try <- seq(0, 500e-6, length.out = 10000)
carbs <- carbAddition(dic_init = dic_init,
                    ta_init = ta_init,
                    species = "co3",
                    mineral_addition = CO3_try,
                    S = S,
                    T = T) %>%
  bind_cols(data.frame(CO3_addition = CO3_try))

# Find the value of co3 that minimizes the difference
carbs %>%
  filter(pH == carbs$pH[which.min(abs(pHFinal - carbs$pH))]) %>%
  select(pH:OmegaAragonite, CO3_addition) -> carbs_5mm

# Print the amount of carbonate added
cat("Amount of CO3²⁻ added:", carbs_5mm$CO3_addition*1e6, "μmol/kg\n")

# Calculate the final parameters
dic_co3 <- dic_init + carbs_5mm$CO3_addition[1]
ta_co3 <- ta_init + 2*carbs_5mm$CO3_addition[1]


# Find what oh added would get us to pH
oh_try <- seq(0, 500e-6, length.out = 10000)
carbs <- carbAddition(dic_init = dic_init,
                    ta_init = ta_init,
                    species = "oh",
                    mineral_addition = oh_try,
                    S = S,
                    T = T) %>%
  bind_cols(data.frame(oh_addition = oh_try))

# Find the value of co3 that minimizes the difference
carbs %>%
  filter(pH == carbs$pH[which.min(abs(pHFinal - carbs$pH))]) %>%
  select(pH:OmegaAragonite, oh_addition) -> carbs_5mm

# Print the amount of carbonate added
cat("Amount of OH⁻ added:", carbs_5mm$CO3_addition*1e6, "μmol/kg\n")

# Calculate the final parameters
dic_oh <- dic_init
ta_oh <- ta_init + carbs_5mm$oh_addition[1]



# Find the value of co3 that minimizes the difference
# carbs %>%
#   filter(pH == carbs$pH[which.min(abs(pHFinal[2] - carbs$pH))]) %>%
#   select(pH:OmegaAragonite, CO3_addition) -> carbs_15mm
# 
# # Print the amount of carbonate added
# cat("Amount of CO3²⁻ added:",carbs_15mm$CO3_addition*1e6, "μmol/kg\n")
# 
# # Calculate the final parameters
# dic_final <- dic_init + carbs_15mm$CO3_addition[1]
# ta_final <- ta_init + 2*carbs_15mm$CO3_addition[1]

```

Chan & Connolly (2013) meta analysis suggests that there is an approximately ~15% change in calcification rates per unit change in omega. This range is only valid over omega 2-4.

```{r}
omega_init = carb(flag = 15,
                  var1 = ta_init,
                  var2 = dic_init, 
                  S = S, T = T)

omega_co3 = carb(flag = 15,
                  var1 = ta_co3,
                  var2 = dic_co3, 
                  S = S, T = T)

omega_oh = carb(flag = 15,
                  var1 = ta_oh,
                  var2 = dic_oh, 
                  S = S, T = T)

cat("Initial omega calculated:", omega_init, "\n")
cat("Final omega calculated assuming CO3:", omega_co3, "\n")
cat("Final omega calculated assuming OH:", omega_oh, "\n")
cat("Difference omega calculated:", c((omega_co3 - omega_init),(omega_oh - omega_init)), "\n")
cat("Predicted delta G based on omega difference:", (omega_oh$OmegaAragonite - omega_init$OmegaAragonite)*0.15, "\n")



# langdon 2000 calcs
k = c(1.2e7, 0.9e7,1.5e7) # lower = 0.9, upper = 1.5
n = c(0.69, 0.65, 0.73) # lower
C = c(-89, -93, -85)
Ca = 10.12e-3

G_init = k*(Ca^n)*omega_init$CO3+C
G_oh = k*(Ca^n)*omega_oh$CO3+C

(G_oh - G_init) / G_init * 100 # 92% predicted increase mean
```
This underestimation is to be expected due to the nonlinear effect of omega on G (Anthony et al. 2011), and the per unit reduction range is wide (0-31% per unit of omega decreae, 95% CI), granting an upper predicted G increase of 37%. Our G increaes (~48%) are in line with studies that tested a larger range of aragonites including Langdon et al. (2000), Ohde & Hussain (2004)

# Property-Property Plots

```{r}
library(metR)
library(geomtextpath)

# Define range for DIC and TA
DIC_range <- seq(2050, 2300, by = 5) # umol/kg
TA_range <- seq(2350, 2650, by = 5)  # umol/kg

# Generate grid of DIC and TA
dat <- expand.grid(DIC = DIC_range, TA = TA_range) %>%
# Calculate carbonate chemistry using seacarb (flag=15 means DIC & TA given)
# Assume salinity = 35, temperature = 25°C, pressure = 0 dbar
  mutate(carb(
  flag = 15,
  var1 = TA/10^6,
  var2 = DIC/10^6,
  S = 34, T = 27, P = 0, Patm = 1, Pt = 0, Sit = 0
)) %>%
  select(DIC, TA, pH, pCO2, OmegaAragonite) %>%
  mutate(DIC = DIC*10^6,
         alpha = case_when(abs(pH - 8.16) <= 0.02 ~ 1.0,
                           TRUE ~ 0.5))

carb_points <- data.frame(name = c('experimental', 'co3 addition', 'oh addition'),
           DIC = c(2108, 2201.959, 2108),
           TA = c(2436, 2623.919, 2517.108))

# Plot using ggplot
ggplot(dat, aes(x = DIC, y = TA, fill = pH, z=pH)) +
  geom_raster(interpolate = TRUE) +
  scale_fill_gradientn(colours = paletteer::paletteer_c("grDevices::rainbow", 7)) + # 
  geom_contour(color = "black", alpha = 0.7, linewidth = 0.6,
               binwidth =  0.1) +
  geom_textcontour(binwidth = 0.1, size = 3, spacing=0.5, gap=TRUE,
                   vjust=0, hjust=0.4) +
  geom_contour_filled(aes(z = pH),
                        breaks = c(8.14, 8.18),
                        show.legend = FALSE,
                        alpha = 0.0,
                        linetype = "dashed",
                        color = "black",
                        fill = NA,
                        linewidth = 1) +
  # geom_contour(aes(x=DIC, y=TA, z=OmegaAragonite),
  #              data = subset(dat, pH>=8.14 & pH<=8.18),
  #             color = "black", alpha = 0.7, linewidth = 0.5,
  #              binwidth = 0.2,
  #              inherit.aes = F) +
  # geom_textcontour(aes(x=DIC, y=TA, z=OmegaAragonite),
  #              data = subset(dat, pH>=8.14 & pH<=8.18),
  #              binwidth = 0.2,
  #              size = 3, gap=FALSE,
  #              spacing = 100,
  #              vjust=0) +
  geom_point(data = carb_points, aes(DIC, TA),
             color="darkred", size=3, inherit.aes = F) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = expression("Dissolved Inorganic Carbon ("*mu*"mol kg"^{-1}*")"),
       y = expression("Total Alkalinity ("*mu*"mol kg"^{-1}*")"),
       fill = "pH") +
  annotate("segment",
           x = 2115, y = 2445,
           xend = 2198, yend = 2609,
           arrow = arrow(length = unit(0.2, "cm")),
           color = "darkred", linewidth = 2) +
  # annotate("text", label = expression("CO"[3]^{"2-"}),
  #    x = 2150, y = 2540, fontface="bold",
  #    color = "darkred", size = 5, angle=45) +
  annotate("segment",
         x = 2108, y = 2444,
         xend = 2108, yend = 2506,
         arrow = arrow(length = unit(0.2, "cm")),
         color = "darkred", linewidth = 2) +
  # annotate("text", label = bquote("OH"^"-"),
  #      x = 2100, y = 2471, fontface="bold",
  #      color = "darkred", size = 5, angle=90) +
  theme_minimal()

ggsave('../figures/property-propertyPlot.tiff',
       dpi = 600,
       width = 6,
       height = 5,
       units = "in",
       device = "tiff",
       compression = "lzw")
```

